Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''

### 2019-05-29 second-draft/pnas-word-order-si.tex

\textcolor{red}{parseability vs parsability}

@@

@@

Greater values of $\lambda$ can mathematically result in degenerate solutions.
To show this, note that the following inequality always holds:
\begin{equation}
I[U; T] \leq H[U].
\end{equation}
Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
\begin{equation}
    R_{Eff} = I[U; T] - \lambda H[U] \leq 0.
\end{equation}
and it takes the maximally possible value zero if $U$ maps all trees to a single utterance, in which case both $I[U; T]$ and $H[U]$ are zero.
This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parseability, whereas values closer to $0$ diminish the role of predictability.

@@

@@

In this section, we describe how we optimized grammar parameters for efficiency.
A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $\mathcal{L}_\theta(t) = \mathbf{w}$ linearizing the tree.
More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ defines a \emph{probability distribution} $p_{\mathcal{L}_\theta}(u|t)$ over ordered sequences of words $\mathbf{w}$.
In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $p_{\mathcal{L}_\theta}(u|t)$ concentrates on a single ordering $u$.

Recall the definition of efficiency
\begin{equation}\label{eq:efficiency-recall}
	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
\end{equation}
where
\begin{equation}
	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
\end{equation}
\begin{equation}
	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
\end{equation}
where $t \sim \mathcal{T}$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.

@@

%	R_{Pars}^\phi := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)

@@

%The parsers and language models appropriate to a tree distribution $\mathcal{T}$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:

@@

At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
Empirically, we observe that optimizing differentiable ordering grammars for efficiency leads to convergence towards deterministic behavior, allowing us to extract equivalent deterministic grammars as described in Section~\ref{sec:diff-gramm}.

@@

We used the CKY algorithm to compute the parsing loss $H[T|\mathcal{L}_\theta(T)]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
% TODO get UAS for compatibility with other parts of SI
%We then compared parseability of random and optimized word order grammars on English data.

\begin{figure}
    \centering
    \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
	\caption{Parsing loss $H[T|\mathcal{L}_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue). We report $H[T|\mathcal{L}_\theta(T)]$ normalized by sentence length.}
    \label{fig:cky-parser}
\end{figure}

@@

    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[T|L_\theta(T)]$ (lower is better), for the four distorted orderings, and the actual orderings (`real'). We report $H[T|\mathcal{L}_\theta(T)]$ normalized by sentence length.}
    \label{fig:distorted-parser}
\end{figure}

@@

        \caption{Parsing loss ($H[T|\mathcal{L}_\theta(T)]$, normalized by sentence length) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
    \label{fig:learning-czech}
\end{figure}

>>>>

\textcolor{red}{word limit}

\textcolor{red}{grammar example in SI}

\textcolor{red}{cover letter goes to broad person}
\textcolor{red}{parseability vs parsability}

@@

\textcolor{red}{Czech parser}

\textcolor{red}{random all 50}

@@

Greater values of $\lambda$ can mathematically result in degenerate solutions.
To show this, note that the following inequality always holds:
\begin{equation}
I[\utterance; \tree] \leq H[\utterance].
\end{equation}
Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
\begin{equation}
    R_{Eff} = I[\utterance; \tree] - \lambda H[\utterance] \leq 0.
\end{equation}
and it takes the maximally possible value zero if there is only a single utterance $\utterance$, in which case both $I[\utterance; \tree]$ and $H[\utterance]$ are zero.
This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parseability, whereas values closer to $0$ diminish the role of predictability.

@@

\paragraph{Grammar Example}

\begin{tabular}{llllllllll}
	                             & \multicolumn{3}{English}    &   \\
	Relation                     &     & $a_\tau$ & $b_\tau$   &   \\
	object (\textit{obj})        &     &          &            &   \\
	oblique (\textit{obl})       &     &          &            &   \\
	case (\textit{lifted\_case}) &     &          &            &   \\
\end{tabular}

@@

In this section, we describe how we optimized grammar parameters for efficiency.
A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $u = \mathcal{L}_\theta(t)$ linearizing the tree.
More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ defines a \emph{probability distribution} $p_{\mathcal{L}_\theta}(u|t)$ over ordered sequences of words $u$.
In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $p_{\mathcal{L}_\theta}(u|t)$ concentrates on a single ordering $u$.

Recall the definition of efficiency
\begin{equation}\label{eq:efficiency-recall}
	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
\end{equation}
where
\begin{equation}
	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
\end{equation}
\begin{equation}
	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
\end{equation}
where $t \sim \tree$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.

@@

%	R_{Pars}^\phi := I[\utterance,\tree] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)

@@

%The parsers and language models appropriate to a tree distribution $\tree$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:

@@

At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
The neural parser and language model are also randomly initialized at the beginning of optimization.
Empirically, we observe that optimizing differentiable ordering grammars for efficiency leads to convergence towards deterministic behavior, allowing us to extract equivalent deterministic grammars as described in Section~\ref{sec:diff-gramm}.

@@

We used the CKY algorithm to compute the syntactic ambiguity $H[\tree|\utterance]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
% TODO get UAS for compatibility with other parts of SI
%We then compared parseability of random and optimized word order grammars on English data.

\begin{figure}
    \centering
    \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
	\caption{Parsing loss $H[\tree|\utterance]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue). We report $H[\tree|\utterance]$ normalized by sentence length.}
    \label{fig:cky-parser}
\end{figure}

@@

    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[\tree|\utterance]$ (lower is better), for the four distorted orderings, and the actual orderings (`real'). We report $H[\tree|\utterance]$ normalized by sentence length.}
    \label{fig:distorted-parser}
\end{figure}

@@

	\caption{Parsing loss ($H[\tree|\utterance]$, normalized by sentence length) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
    \label{fig:learning-czech}
\end{figure}
